### 2022-07-11 paper_ellis/abstract.tex
rename paper/ to paper_ellis/

We introduce the college application problem, a submodular optimization problem involving the maximization of the expected maximum value of a portfolio of random variables subject to a budget constraint. We prove that this problem is NP-complete and provide two optimization algorithms based on dynamic programming. A modification of the second dynamic program yields a fully polynomial-time approximation scheme. These results suggest that the college application problem is a relatively easy instance of submodular maximization.