### 2022-07-14 paper_alma/abstract.tex
refactor paper_alma to match paper_ellis, expand introduction

This paper considers a novel portfolio optimization problem called the college application problem. We show that the objective function is a nondecreasing submodular set function and provide a quadratic-time solution algorithm. We discuss the properties of the optimal solution and their implications for student welfare.