### 2022-07-18 paper_alma/body.tex
rm linearization heuristic; begin refactor of discussion

\section{Introduction}

In competitive college admissions markets, the optimal college application strategy is a matter of perennial interest due to the widespread perception that future career earnings depend on college outcomes and the challenge of balancing risk and reward across the schools in one's application portfolio. Research in the social sciences has shown that students differ in their application styles: Students from wealthy backgrounds tend to apply to many colleges, including several so-called ``reach schools'' where their probability of admission is relatively low, whereas low-income students tend to apply to just a few schools, and favor those where admission is a safer bet \cite{hartlaubandschneider2012,vanhuizenandalessie2019}. Does this phenomenon reflect systematic differences in students' risk aversion, or rational decisionmaking under different resource constraints?

Due to the widespread perception that college admissions results determine future career outcomes, considerable demand has arisen for private consulting services that help students decide where to apply. In the American college consulting industry, students pay private consultants an average of \$200 per hour for assistance in preparing application materials, estimating their admissions odds, and identifying target schools \cite{sklarow2018}. In Korea, an important revenue stream for admissions consulting firms such as Megastudy (\url{megastudy.net}) and Jinhak (\url{jinhak.com}) is ``mock application'' software that claims to use artificial intelligence to optimize the client's application strategy.

Despite the size of the market for prescriptive admissions counseling, most quantitative research on college admissions to date has been descriptive in nature. A standard benchmark for logistic regression techniques, for example, involves estimating students' admissions probabilities from information about their past academic performance \cite{acharyaetal2019,lim2013}. When professional admissions consultants advise their clients on where to apply to college, however, they typically rely on qualitative heuristics, such as balancing applications between safety and reach schools, rather than optimization models \cite{jeon2015,peck2021}. (Admissions consultants refer to schools with low utility and high admissions probability as \emph{safety schools}, and those with high utility and low admissions probability as \emph{reach schools.}) How effective are these heuristics, and can we improve them?

In the present study, we formulate the student's college application decision as the following combinatorial optimization problem:
\begin{align} \label{headlineproblem}
\begin{split}
\text{maximize}\quad & v(\mathcal{X}) =  \operatorname{E}\Bigl[\max\bigr\{t_0,
\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
\text{subject to}\quad & \mathcal{X} \subseteq \mathcal{C}, ~~|\mathcal{X}| \leq H
\end{split}
\end{align}
Here $\mathcal{C} = \{ 1 \dots m\}$ represents the set of colleges in the market and $h > 0$ is the number of colleges the student can apply to. For $j = 1 \dots m$, let $f_j$ denote the student's probability of being admitted to school $j$ if she applies and $Z_j$ a random, independent Bernoulli variable with probability $f_j$, so that $Z_j$ equals one if she is admitted and zero if not. It is appropriate to assume that the $Z_j$ are statistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates). For $j = 0\dots m$, $t_j\geq 0$ indicates the utility she associates with attending school $j$, where her utility is $t_0$ if she does not attend college. Then the student's objective is to maximize the expected utility associated with the best school she is admitted to. Therefore, her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which she applies, which we call her application \emph{portfolio.}

The problem can also be applied to other competitive matching games such as job application or marriage proposal. Here the budget constraint may arise due to the time required to communicate with each applicant or a legal limit on the number of applications allowed.

We show that the optimal portfolios for the cardinality-constrained college are \emph{nested} in the budget constraint, meaning that when $\mathcal{X}_h$ is the optimal portfolio of size $h$, $\mathcal{X}_h \subset \mathcal{X}_{h+1}$. The nestedness property is equivalent to the optimality of the \emph{greedy} algorithm that iteratively adds to $\mathcal{X}$ the school that elicits the greatest increase in the objective value. The proof of the nestedness property yields insight into the properties of the optimal college application strategy: Students who cannot afford to apply to many schools allocate a greater proportion of their application budget to safety schools than students with a large application budget. This phenomenon, which we call optimal risk elasticity, reveals a shortcoming of the reach/safety school heuristics used in the admissions consulting industry: Whereas admissions consultants recommend a \emph{static} risk allocation that allocates a fixed proportion of the portfolio to reach, target, and safety schools, we find that the optimal application strategy gradually favors reach schools over safety schools as $h$ increases. Optimal risk elasticity therefore provides a \emph{rational} explanation for the divergent risk behavior observed in the social-science research cited above: It may not be the case that low-income students are inherently more risk-averse than their wealthy peers, but rather, their constrained budget induces a more risk-averse optimal policy.

Our analysis also provides a tight, constant upper bound on the marginal value of additional applications, which may be informative in the design of fee waiver policies and setting application limits in centralized admissions markets.
% \subsection{Literature review}

\subsection{Related optimization problems}

The objective function is a nondecreasing, submodular set function in the sense first described by Nemhauser et al. \cite{nemhauseretal1978}. They showed that the greedy algorithm is asymptotically $(1 - 1/e)$-optimal for optimizing a monotonic submodular function over a cardinality constraint, and a subsequent result of Nemhauser and Wolsey \cite{nemhauserandwolsey1978} proved that this approximation ratio is the best achievable by a polynomial-time algorithm. In the special case of college application, however, we show that the same algorithm is exact.

% define nestedness
The proof of the optimality of the greedy algorithm rests on the nested structure of the portfolios, a property that is of independent interest. As Rozanov and Tamir \cite{rozanovandtamir2020} articulate, the knowledge that the optima are nested aids not only in computing the optimal solution, but in the implementation thereof under uncertain information. For example, in the United States, many college applications are due at the beginning of November, and it is typical for students to begin working on their applications during the prior summer because colleges reward students who tailor their essays to the target school. However, students may not know how many schools they can afford to apply to until late October. The nestedness property---or equivalently, the validity of a greedy algorithm---implies that even in the absence of complete budget information, students can begin to carry out the optimal application strategy by writing essays for schools in the order that they enter the optimal portfolio.
% Greedy algorithms produce a \emph{nested} family of solutions parameterized by the budget $H$: If $H \leq H'$, then the greedy solution for budget $H$ is a subset of the greedy solution for budget $H'$.

% Reword next two paragraphs to prevent collision with Ellis paper. emphasize cardinality over knapsack
The college application problem can be viewed as a static variant of the Pandora's Box problem proposed by Weitzman \cite{weitzman1979}. In a Pandora's Box formulation of college application, the student applies to schools one by one, each time paying an application fee and observing her admissions outcome after a certain time delay. The problem is to determine an optimal \emph{stopping rule} for when Pandora should halt her college search and accept the best admissions offer she has on hand. Weitzman showed that the optimal policy is to stop searching when the value of the current best offer exceeds the maximum \emph{reservation price,} a statistic that represents the expected value of applying to a new college.

Arguably, the static model considered in this study is more hostile to students than the Pandora's Box problem. If, for example, an unlucky Pandora is rejected from a safety school at an early round of application, then she can compensate for the unexpected loss by pivoting to a more risk-averse application strategy. By contrast, the decisionmaker in our college application problem must commit at the outset to applying to every school in her application portfolio. The admissions process used in the United States can be viewed as the concatenation of both problems: In the fall, students solve \eqref{headlineproblem} and send out a batch of applications. Then, upon observing their admissions outcomes in March, they use the Pandora strategy to pursue additional offers by applying to schools that offer rolling admissions.

\subsection{Admissions markets}

%Fu \cite{fu2014} considered a problem that is reducible to ours: since application is at cost rather than a constraint, the objective function is separable in the cost and therefore must be optimal over portfolios of the given size. Thus, solving her problem reduces to solving $m$ instances of \eqref{headlineproblem}.

% Merge into above
Some economists have constructed equilibrium models of the college admissions process that feature the student's application decision as a prominent subproblem. Fu \cite{fu2014}, for example, modeled the United States admissions market as a sequential game played by colleges and students. Colleges announce tuition rates, students apply to college, colleges announce admissions decisions and financial aid offers, and lastly, students decide where to enroll. Equilibrium arises when no college can improve its expected utility by modifying its tuition or admissions policy, and no student can improve her expected utility by modifying her application strategy. The ultimate goal of Fu's model was a comparative statics analysis of various hypothetical reforms to the structure of the admissions market; hence, it was adequate to cluster colleges into $m=8$ broad categories. Therefore, though estimating Fu's model required solving a problem similar to \eqref{headlineproblem}, it was a small instance that could be solved by enumerating all possible portfolios. Our study pursues a more general solution.

The literature on admissions markets also includes a vast body of research on the deferred acceptance algorithm, an algorithm for matching students to schools having finite capacities that generalizes the Gale--Shapley algorithm for stable marriage. The algorithm takes students' ordinal preferences over schools and schools' ordinal preferences over students as input, and produces an assignment that possesses a number of desirable properties: It is a stable matching, meaning that no student--school pair is incentivized to deviate from the assignment. And the preference-reporting mechanism is incentive compatible, meaning that no student can improve her outcome by lying about her preferences \cite{galeandshapley1962,roth1982}. In design, deferred acceptance is a centralized assignment algorithm; however, stable assignment also be interpreted as the equilibrium of a decentralized admissions game in which agents have perfect information about one another's preferences \cite{azevedoandleshno2016}.

The utility model implied by stable assignment differs from that of the present study in two key ways: First, under stable assignment, students' preferences are ordinal rather than cardinal, although it is possible to estimate cardinal utility values from stable matchings in certain special cases, such as when colleges' preferences are identical and students' preferences are determined by the multinomial logit choice model \cite{kapur2021}. Second, stable assignment is deterministic: The typical student-proposing deferred acceptance algorithm always produces the unique, student-optimal stable assignment. There exist variants of the deferred acceptance algorithm that introduce randomization in order to break ties in schools' preference lists or optimize for distributional goals such as gender parity, but because the randomization occurs after students submit their preferences, it has no effect on the optimal application strategy, which remains to report one's preferences honestly \cite{ashlagiandnikzad2020,bodohcreed2020}. In the admissions process considered in this study, students face a tougher strategic challenge.

To the best of our knowledge, the first systematic study of the college application problem was undertaken by Kapur in his master's thesis \cite{kapur2022}, which considers \eqref{headlineproblem} as well as the more general case of college application with a knapsack constraint. The present study extends Kapur's results with a further analysis of the accuracy of the heuristics used in the admissions consulting industry and the implications of the nestedness and submodularity properties for student welfare.

\subsection{Outline}

This paper has five sections. In section \ref{sectionModel}, we derive a closed-form expression for the objective function of \eqref{headlineproblem}, show that it is a monotonic submodular function, and state a technical lemma. In section \ref{sectionGreedy}, we prove that the optimal portfolios are nested in the budget $h$, which implies the validity of the greedy algorithm. Further refinement reduces the computation time from $O(h^2 m + m\log m)$ to $O(hm)$. In section \ref{sectionDiscussion}, we examine alternative solution mechanisms used in the admissions consulting industry, their performance relative to the optimal solution, and the implications for student welfare. A brief conclusion follows in section \ref{sectionConclusion}.

\section{Preliminaries} \label{sectionModel}

For the remainder of the paper, unless otherwise noted, we assume with trivial loss of generality that each $f_j \in (0, 1]$ and $t_0 < t_1 \leq \cdots \leq t_m$. Unless otherwise noted, we assume $t_0 = 0$, an assumption justified presently. 
% Lots of overlap here with paper_ellis; edit there first and then import

\subsection{The objective function} \label{sectionObjective}

First we derive a closed-form expression for the objective function of \eqref{headlineproblem}.

We refer to the set $\mathcal{X} \subseteq \mathcal{C}$ of schools to which a student applies as her \emph{application portfolio.} The expected utility the student receives from $\mathcal{X}$ is called its \emph{valuation}. Given an application portfolio, let $p_j(\mathcal{X})$ denote the probability that the student attends school $j$. This occurs if and only if she \emph{applies} to school $j$, is \emph{admitted} to school $j$, and is \emph{rejected} from any school she prefers to $j$; that is, any school with higher index. Hence, for $j= 0\dots m$,
\begin{align}
p_j(\mathcal{X}) &= 
\begin{cases}
\displaystyle f_j  \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}), \quad & j \in \{0\}\cup\mathcal{X}\\
0, \quad & \text{otherwise}
\end{cases} 
\end{align}
The following proposition follows by computing $v(\mathcal{X}) = \sum_{j=0}^m  t_j p_j(\mathcal{X})$.
\begin{proposition}[Closed form of portfolio valuation function]
\begin{align}
v(\mathcal{X}) &= \sum_{j=0}^m t_j p_j(\mathcal{X}) = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
%\qquad v(x) &= t_0 \prod_{j=1}^m (1 - f_{j} x_j) + \sum_{j=1}^m \Bigl( x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \Bigr) \label{closedformportfoliovaluationx}
\end{align}
\end{proposition}
%\begin{proof}Computing $v(\mathcal{X}) = \sum_{j=0}^m  t_j p_j(\mathcal{X})$ yields \eqref{closedformportfoliovaluationX}. Next, because $1 - f_j x_j = 1$ if $x_j = 0$, we may define $p_j$ equivalently as $p_j(x) = x_j  f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’})$ to obtain \eqref{closedformportfoliovaluationx}. 
%\end{proof}

Next, we show that without loss of generality, we may assume that $t_0 = 0$.

\begin{lemma} \label{assumetzerozero}
For some $\gamma \leq t_0$, let $\bar t_j = t_j - \gamma$ for $j = 0 \dots m$. Then $v(\mathcal{X}; \bar t_j) = v(\mathcal{X};  t_j) -  \gamma$ for any $\mathcal{X} \subseteq \mathcal{C}$. 
\end{lemma}
\begin{proof}
By definition, $\sum_{j=0}^m p_j(\mathcal{X}) = \sum_{j \in \{0\}\cup\mathcal{X}} p_j(\mathcal{X}) = 1$. Therefore
\begin{align}
\begin{split}
v(\mathcal{X}; \bar t_j) &= \sum_{j\in \{0\}\cup\mathcal{X}}  \bar t_j p_j(\mathcal{X})
=\sum_{j\in \{0\}\cup\mathcal{X}} (t_j - \gamma) p_j(\mathcal{X}) \\
&=\sum_{j\in \{0\}\cup\mathcal{X}} t_j p_j(\mathcal{X})  - \gamma 
= v(\mathcal{X}; t_j) - \gamma
\end{split} 
\end{align}
which completes the proof.
\end{proof}

\subsection{An elimination technique} \label{eliminationtechniquesection}

Now, we present a variable-elimination technique that will prove useful throughout the paper.\footnote{We thank Yim Seho for pointing out this useful transformation.} Suppose that the student has already resolved to apply to school $k$, and the remainder of her decision consists of determining which \emph{other} schools to apply to. Writing her total application portfolio as $\mathcal{X} = \mathcal{Y} \cup \{k\}$, we devise a function $w(\mathcal{Y})$ that orders portfolios according to how well they ``complement'' the singleton portfolio $\{k\}$. Specifically, the difference between $v(\mathcal{Y} \cup\{k\})$ and $w(\mathcal{Y})$ is the constant $f_k t_k$.

To construct $w(\mathcal{Y})$, let $\tilde t_j$ denote the expected utility the student receives from school $j$ \emph{given} that she has been admitted to school $j$ and applied to school $k$. For $j < k$ (including $j = 0$), this is $\tilde t_j = (1- f_k) t_j + f_k t_k$; for $j > k $, this is $\tilde t_j = t_j$. This means that 
\begin{equation}\label{Vyastildet}
v(\mathcal{Y}\cup\{k\}) = \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y}).\end{equation}
The transformation to $\tilde t$ does not change the order of the $t_j$-values. Therefore, the expression on the right side of \eqref{Vyastildet} is itself a portfolio valuation function. In the corresponding market, $t$ is replaced by $\tilde t$ and $\mathcal{C}$ is replaced by $\mathcal{C}\setminus\{k\}$. To restore our convention that $t_0 = 0$, we obtain $w(\mathcal{Y})$ by taking $\bar t_j = \tilde t_j - \tilde t_0$ for all $j \neq k$ and letting
\begin{equation}  \label{wYvXminusconst}
w(\mathcal{Y})
= \sum_{j \in \{0\} \cup \mathcal{Y}} \bar t_j p_j(\mathcal{Y})
= \sum_{j \in \{0\} \cup \mathcal{Y}} \tilde t_j p_j(\mathcal{Y})- \tilde t_0
= v(\mathcal{Y}\cup\{k\}) -  f_k t_k \end{equation}
where the second equality follows from Lemma \ref{assumetzerozero}. The validity of this transformation is summarized in the following theorem, where we write $v(\mathcal{X}; \bar t)$ instead of $w(\mathcal{Y})$ to emphasize that $w(\mathcal{Y})$ is, in form, a portfolio valuation function.

\begin{lemma}[Eliminate school $k$] \label{eliminationtheorem}
For $\mathcal{X} \subseteq \mathcal{C} \setminus \{k\}$, $v(\mathcal{X}\cup\{k\}; t)  = v(\mathcal{X}; \bar t) + f_k t_k$, where
\begin{align}\label{howtotransformtj}
\bar t_j = 
\begin{cases}
(1 - f_k) t_j, \quad & t_j \leq t_k \\
t_j - f_k t_k, \quad& t_j > t_k.
\end{cases}
\end{align}
\end{lemma}

\begin{proof}
It is easy to verify that \eqref{howtotransformtj} is the composition of the two transformations (from $t$ to $\tilde t$, and from $\tilde t$ to $\bar t$) discussed above.
\end{proof}

%\noindent The transformation \eqref{howtotransformtj} can be applied iteratively to accommodate the case where the student has already resolved to apply to multiple schools.

\subsection{Submodularity of the objective}

Now, we show that the portfolio valuation function is submodular. This result is primarily of taxonomical interest and may be safely skipped, as our subsequent results do not rely on submodular analysis.

\begin{definition}[Submodular set function]
Given a ground set $\mathcal{C}$ and function $v : 2^{\mathcal{C}} \mapsto \mathbb{R}$, $v(\mathcal{X})$ is called a \emph{submodular set function} if and only if $v(\mathcal{X}) + v(\mathcal{Y}) \geq v(\mathcal{X}\cup\mathcal{Y}) + v(\mathcal{X}\cap\mathcal{Y})$
for all $\mathcal{X}, \mathcal{Y} \subseteq \mathcal{C}$. Furthermore, if $ v(\mathcal{X}\cup\{k\}) - v(\mathcal{X}) \geq 0$ for all $\mathcal{X} \subset \mathcal{C}$ and $k \in \mathcal{C} \setminus \mathcal{X}$, $v(\mathcal{X})$ is said to be a \emph{nondecreasing} submodular set function.
\end{definition}

\begin{theorem}
The college application portfolio valuation function
$v(\mathcal{X})$ % = \sum_{j\in\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\]
is a nondecreasing submodular set function.
\end{theorem}

\begin{proof}
It is self-evident that $v(\mathcal{X})$ is nondecreasing. To establish its submodularity, we apply proposition 2.1.iii of \cite{nemhauserandwolsey1978} and show that
\begin{equation}\label{nemhauseriii}
v(\mathcal{X} \cup \{j\}) - v(\mathcal{X}) \geq 
v(\mathcal{X} \cup \{j, k\}) - v(\mathcal{X} \cup \{k\})
\end{equation}
for $\mathcal{X} \subset \mathcal{C}$ and $j \neq k \in \mathcal{C} \setminus \mathcal{X}$. By Lemma \ref{eliminationtheorem}, we can repeatedly eliminate the schools in $\mathcal{X}$ according to \eqref{howtotransformtj} to obtain a portfolio valuation function $w(\mathcal{Y})$ with parameter $\bar t$ such that $w(\mathcal{Y}) = v(\mathcal{X} \cup \mathcal{Y}) + \text{const.}$ for any $\mathcal{Y} \subseteq \mathcal{C} \setminus \mathcal{X}$. Therefore, \eqref{nemhauseriii} is equivalent to
\begin{align}
& w(\{j\}) - w(\varnothing) \geq w(\{j, k\}) - w(\{k\}) \\
\iff \qquad &w(\{j\})  +  w(\{k\})  \geq w(\{j, k\})  \\
\iff \qquad &\operatorname{E}[\,\bar t_j Z_j\,] + \operatorname{E}[\,\bar t_k Z_k\,] 
\geq \operatorname{E}\bigl[\max\{ \bar t_j Z_j, \bar t_k Z_k \} \bigr]
\end{align}
which is true because $t_j$ and $t_k$ are nonnegative.
\end{proof}

\section{The greedy algorithm} \label{sectionGreedy}

\subsection{The nestedness property} 
The optimality of the greedy algorithm for the college application problem rests on the fact the the optimal solution possesses a special structure: An optimal portfolio of size $h+1$ includes an optimal portfolio of size $h$ as a subset.

\begin{theorem}[Nestedness of optimal application portfolios] \label{nestedapplication}
There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ satisfying the nestedness relation

\begin{equation}
\mathcal{X}_1 \subset \mathcal{X}_2\subset \dots \subset \mathcal{X}_m
\end{equation}
such that each $\mathcal{X}_h$ is an optimal application portfolio when the application limit is $h$.
\end{theorem}

\begin{proof}
By induction on $h$. Applying Lemma \ref{assumetzerozero}, we assume that $t_0 = 0$.

(Base case.) First, we will show that $\mathcal{X}_1 \subset \mathcal{X}_2$. To get a contradiction, suppose that the optima are $\mathcal{X}_1 = \{j\}$ and $\mathcal{X}_2 = \{k, l\}$, where we may assume that $t_k \leq t_l$. Optimality requires that

\begin{equation}v(\mathcal{X}_1 )  = f_j t_j > v(\{k\}) = f_k t_k\end{equation}
and
\begin{align}
\begin{split}
v(\mathcal{X}_2) =  f_k (1- f_l) t_k + f_l t_l &> v(\{j, l\}) \\
& = f_j (1- f_l) t_j + (1- f_j) f_l t_l + f_j f_l \max\{t_j, t_l\} \\
&\geq  f_j (1- f_l) t_j + (1- f_j) f_l t_l + f_j f_l  t_l \\
&= f_j (1- f_l) t_j + f_l t_l  \\
&\geq f_k (1- f_l) t_k + f_l t_l  = v(\mathcal{X}_2)
\end{split}
\end{align}
which is a contradiction.

(Inductive step.) Assume that $\mathcal{X}_1 \subset \cdots \subset \mathcal{X}_h$, and we will show $\mathcal{X}_h \subset \mathcal{X}_{h+1}$. Let $k = \argmax\{ t_k: k \in \mathcal{X}_{h+1}\}$ and write $\mathcal{X}_{h+1} = \mathcal{Y}_{h} \cup \{k\}$.

Suppose $k \notin \mathcal{X}_h$. To get a contradiction, suppose that $v(\mathcal{Y}_h) < v(\mathcal{X}_h)$ and  $v(\mathcal{X}_{h+1}) > v(\mathcal{X}_h \cup \{k\})$. Then
\begin{align}
\begin{split}
v(\mathcal{X}_{h+1})&= v(\mathcal{Y}_{h} \cup \{k\}) \\
&= (1 - f_k) v(\mathcal{Y}_h) + f_k t_k \\
&\leq (1 - f_k) v(\mathcal{X}_h) + f_k \operatorname{E}\bigl[ \max\{t_k, X_h\}\bigr]\\
&=  v(\mathcal{X}_h\cup \{k\})
\end{split}
\end{align}
is a contradiction.

Now suppose that $k \in \mathcal{X}_h$. We can write $\mathcal{X}_h = \mathcal{Y}_{h-1} \cup \{k\}$, where $ \mathcal{Y}_{h-1}$ is some portfolio of size $h-1$. It suffices to show that $ \mathcal{Y}_{h-1} \subset \mathcal{Y}_h$. By definition, $\mathcal{Y}_{h-1}$ (respectively, $\mathcal{Y}_{h}$) maximizes the function $v(\mathcal{Y}\cup\{k\})$ over portfolios of size $h-1$ (respectively, $h$) that do not include $k$. That is, $\mathcal{Y}_{h-1}$ and $\mathcal{Y}_h$ are the optimal complements to the singleton portfolio $\{k\}$.

Applying Lemma \ref{eliminationtheorem}, we eliminate school $k$, transform the remaining $t_j$-values to $\bar t_j$ according to \eqref{howtotransformtj}, and obtain a function $w(\mathcal{Y}) = v(\mathcal{Y} \cup \{k\}) - f_k t_k$ that grades portfolios $\mathcal{Y} \subseteq \mathcal{C} \setminus \{k\}$ according to how well they complement $\{k\}$. Since $w(\mathcal{Y})$ is itself a portfolio valuation function and $\bar t_0 = 0$, the inductive hypothesis implies that $\mathcal{Y}_{h-1} \subset \mathcal{Y}_h$, which completes the proof.
\end{proof}

\subsection{Quadratic-time solution}

Applying Theorem \ref{nestedapplication} yields an efficient greedy algorithm for the optimal portfolio: Start with the empty set and add schools one at a time, maximizing $v(\mathcal{X}\cup \{k\})$ at each addition. Sorting $t$ is  $O(m \log m)$.  At each of the $h$ iterations, there are $O(m)$ candidates for $k$, and computing $v(\mathcal{X}\cup \{k\})$ is $O(h)$ using \eqref{closedformportfoliovaluationX}; therefore, the time complexity of this algorithm is $O(h^2 m + m \log m)$.

We reduce the computation time to $O(hm)$ by taking advantage of the transformation from Lemma \ref{eliminationtheorem}. Once school $k$ is added to $\mathcal{X}$, we eliminate it from the set $\mathcal{C}\setminus \mathcal{X}$ of candidates, and update the $t_j$-values of the remaining schools according to \eqref{howtotransformtj}. Now, the \emph{next} school added must be the optimal singleton portfolio in the modified market. But the optimal singleton portfolio consists simply of the school with the highest value of $f_j \bar t_j$. Therefore, by updating the $t_j$-values at each iteration according to \eqref{howtotransformtj}, we eliminate the need to compute $v(\mathcal{X})$ entirely. Moreover, this algorithm does not require the schools to be indexed in ascending order by $t_j$, which removes the $O(m\log m)$ sorting cost.

\begin{algorithm}[h] 
%\DontPrintSemicolon
\caption{Optimal portfolio algorithm for the college application problem.} \label{algorithmforlargeh}
\KwIn{Utility values $t \in(0, \infty)^m$, admissions probabilities $f \in (0, 1]^m$, application limit $h \leq m$.}
$\mathcal{C} \gets \{1 \dots m\}$\;
$\mathtt{X, V} \gets $ empty $h$-arrays\;
\For{$i=1\dots h$}
{
    $k \gets \argmax_{j \in \mathcal{C}}\{f_j t_j\}$\;
    $\mathcal{C} \gets \mathcal{C} \setminus \{k\}$\;
    $\mathtt{X}[i] \gets k$\;
     \lIfElse{$i=1$}{$\mathtt{V}[i] \gets f_k t_k$}
     {$\mathtt{V}[i] \gets \mathtt{V}[i-1] + f_k t_k$}
    \For{$j \in \mathcal{C}$}
	{
	\lIfElse{$t_j \leq t_k$}{$t_j \gets (1 -  f_k) t_j $}{$t_j \gets  t_j -  f_k t_k$}
	}
}
\Return{$\mathtt{X, V}$}
\end{algorithm}

\begin{theorem}[Validity of Algorithm \ref{algorithmforlargeh}] \label{validityofalmaalgorithm}
Algorithm \ref{algorithmforlargeh} produces an optimal application portfolio for the cardinality-constrained college application problem in $O(h m)$ time.
\end{theorem}

\begin{proof}
Optimality follows from the proof of Theorem \ref{nestedapplication}. At each of the $h$ iterations of the main loop, finding the top school costs $O(m)$, and the $t_j$-values of the remaining $O(m)$ schools are each updated in unit time. Therefore, the overall time complexity is $O(h m)$.
\end{proof}

By running Algorithm \ref{algorithmforlargeh} with $h = m$, one can obtain the optimal portfolios for \emph{all} $h$ by taking $\mathcal{X}_h = \{ \mathtt{X}[1], \dots, \mathtt{X}[h]\}$. Thus, a trivial modification of the algorithm yields $O(m^2)$-time solution for Fu (2014)'s variant of the college application problem, stated as 
\[ \text{maximize} \quad v(\mathcal{X}) - c(|\mathcal{X}|)\]
where $c(\cdot)$ is an increasing cost function. By separability, the optimal portfolio in this problem must be optimal over portfolios of the same size. Hence, it suffices to compute
\[h^* = \argmax_{h = 1 \dots m} \bigl\{\mathtt{V}[h] - c(h)\bigr\}\] and return $\mathcal{X}_{h^*}$.

\section{Discussion: Heuristics and student welfare} \label{sectionDiscussion}

In this section, we discuss the implications of our model for real-world admissions markets. First, the optimality of the greedy algorithm together with the submodularity of the objective implies that $v(\mathcal{X}_h)$ grows sublinearly in $h$; we provide a tight example for this upper bound. Examining the student profiles for which $v(\mathcal{X}_h)$ grows most rapidly suggests that the underwhelming effectiveness of fee-waiver interventions may be explained by the fact that they target students who have little marginal benefit from additional applications.

Next, we discuss the distributional heuristic, which targets a static risk distribution and (in its inoptimality) reveals that the college application problem possesses a unique property called optimal risk elasticity that is not present in traditional portfolio optimization problems. This property provides alternative explanation for social-science research showing that low-income students use a more risk-averse college application strategy than their wealthy peers.

\subsection{Risk management and student welfare}

Diminishing marginal returns are an intrinsic feature of submodular set functions. In the college application problem, this property, together with the optimality of the greedy algorithm, implies that the gain in student utility from increasing the budget from $h-1$ to $h$ is larger than that from increasing it from $h$ to $h+1$.

\begin{theorem} \label{concavityinh}
For $h = 2 \dots (m-1)$,
\begin{equation}v(\mathcal{X}_h) - v(\mathcal{X}_{h-1}) \geq v(\mathcal{X}_{h+1}) - v(\mathcal{X}_{h}).\end{equation} 
\end{theorem}

\begin{proof}
Applying Theorem \ref{nestedapplication}, we write $\mathcal{X}_h = \mathcal{X}_{h-1} \cup\{j\}$ and $\mathcal{X}_{h+1} = \mathcal{X}_{h-1} \cup\{j, k\}$. By optimality, $v(\mathcal{X}_h) - v(\mathcal{X}_{h-1}) \geq v(\mathcal{X}_{h-1}\cup\{k\}) - v(\mathcal{X}_{h-1})$. By submodularity and nestedness, $v(\mathcal{X}_{h-1}\cup\{k\}) - v(\mathcal{X}_{h-1}) \geq  v(\mathcal{X}_{h}\cup\{k\}) - v(\mathcal{X}_{h}) = v(\mathcal{X}_{h+1}) - v(\mathcal{X}_{h})$.
\end{proof}

\noindent (An elementary proof appears in \cite{kapur2022}.) It follows that when $\mathcal{X}_h$ is the optimal $h$-portfolio for a given market, $v(\mathcal{X}_h) = t_0 + O(h)$. The following example establishes the tightness of this bound.

\begin{example} \label{tightexampleforoneoverhopt}
For a small constant $\varepsilon \in (0, 1)$, define the market as follows.
\begin{center}
\begin{tabular}{r|cccccccc}
$j$   & $1$ & $2$ & $\cdots$ & $m-1$ & $m$ \\ \hline
$f_j$ & $\varepsilon^{1}$ & $\varepsilon^{2}$ & $\cdots$ & $\varepsilon^{m-1}$ & $\varepsilon^{m}$ \\
$t_j$ & $\varepsilon^{-1}$ & $\varepsilon^{-2}$ & $\cdots$ & $\varepsilon^{-(m-1)}$ & $\varepsilon^{-m}$
\end{tabular}%
\end{center}
The optimal portfolios are given by $\mathcal{X}_h = \{m-h + 1, dots, m\}$, and 
\begin{equation*}
v(\mathcal{X}_h) = \sum_{j= m-h +1}^m \Bigl( f_j t_j \prod_{i = j+1}^m (1 - f_{i}) \Bigr) =  \sum_{j= 1}^h  (1 - \varepsilon)^{m-h+j}.
\end{equation*}
Thus, as $\varepsilon$ approaches zero, we have $v(\mathcal{T}_h) \to h$.
\end{example}

The fact that $v(\mathcal{X}_h)$ grows sublinearly in $h$ may provide an economic explanation for the underwhelming efficacy of behavioral nudges, such as application fee waivers, designed to encourage talented low-income students to apply to selective universities \cite{gurantzetal2021}. In our model, the marginal value of an additional college application is greatest to students who have both a small application portfolio, and small admissions odds at the schools they are currently applying to---that is, students whose portolio allocation is skewed toward reach schools. However, fee-waiver interventions have typically targeted high-achieving, low-income students (so-called ``one-offs'' \cite{hoxbyandavery2012,hoxbyandturner2013}) who already have high admissions odds at the schools in their portfolio due a combination of underlying academic talent and the aforementioned tendency of low-income students to favor safety schools. Therefore, our analysis suggests that the students who stand to benefit the most from nudges such as fee waivers are not high-achieving, low-income students, but rather students of middling ability who are prepared for college but unlikely to be admitted to every school.

%On the other hand, dissuading students from applying to too many reach schools may be a desirable feature in an admissions market inasmuch as it eliminates the administrative costs associated with reading low-quality applications. Our analysis of student utility also can inform the design of centralized admissions markets. In the South Korean college admissions process, students are allowed to apply to only $h = 3$ colleges in the main application cycle, a policy which is thought to promote fairness. Suppose that utility is measured in monetary units, $t_0 = \$100$, and $v(\mathcal{X}_3) = \$130$ for a typical student. Then since $v(\mathcal{X}_4) \leq \$140$, a policy that increments the application limit will be unpopular if its tax cost exceeds $\$10$.

\subsection{The distributional heuristic}

In real-world admissions markets, $f_t$ and $t_j$ correlate inversely because desirable schools attract many applicants. In the admissions consulting industry, schools with high utility and low admissions probability are called reach schools, while schools with low utility and high admissions probability are called safety schools, and those in between are called target schools. According to a common heuristic used by admissions consultants, it is best to apply to a roughly equal mix of target, safety, and reach schools \cite{jeon2015,peck2021}. How does this strategy, which we call the \emph{distributional} heuristic, perform compared to the optimal application strategy?

Note that the distributional heuristic does not specify the boundary between target, safety, and reach schools. However, regardless of how this boundary is specified, we argue that the distributional heuristic is limited by the fact that it does not respond adaptively to the application budget $h$. In a market where $f_j$ and $t_j$ are inversely correlated, we find that the first schools to enter the optimal portfolio in the greedy algorithm are safety schools, and the allocation shifts toward to incorporate more target and reach schools as the application budget $h$ grows. In other words, the optimal application strategy gradually favors reach schools over safety schools as $h$ increases.

To see why large values of $h$ favor a risky portfolio allocation, consider two schools $i$ and $j$ having $t_i < t_j$ and $f_i t_i = f_j t_j = 1$. Thus, the actuarial value of applying \emph{individually} to either school is identical, and a risk-neutral student is indifferent between the safety school $i$ and the reach school $j$. Now suppose that the student has resolved to apply to another school, $k$, and the question is whether to apply to  $\mathcal{X} = \{i, k\}$ or $\{j, k\}$. There are three cases:

\begin{itemize}
\item If $t_k \leq t_i < t_j$, then 
\begin{align*}
v(\{i, k\}) -  v(\{j, k\}) &= (1 - f_i) f_k t_k  + f_i t_i - (1 - f_j) f_k t_k - f_j t_j \\
&=  - f_i f_k t_k + f_j f_k t_k &< 0
\end{align*}
because $f_i > f_j$. 
\item If  $t_i \leq t_k \leq t_j$, then 
\begin{align*}
v(\{i, k\}) -  v(\{j, k\}) &= (1 - f_k ) f_i t_i + f_k t_k - (1 - f_j) f_k t_k - f_j t_j \\
&= (1 - f_k) + f_k t_k - \bigl( 1 - \frac{1}{t_j}\bigr) f_k t_k - 1  \\
&=  - f_k + \frac{1}{t_j} f_k t_k & \leq 0
\end{align*}
because $t_k \leq t_j$.
\item If $t_i < t_j \leq t_k$, then 
\begin{align*}
v(\{i, k\}) -  v(\{j, k\}) &=  (1 - f_k) f_i t_i + f_k t_k - (1 - f_k) f_j t_j - f_k t_k -= 0.
\end{align*}
\end{itemize}

\noindent In every case, it is (weakly) optimal to choose the reach school $j$ over the safety school $i$. Thus, however one defines a ``target school,'' it is always better to pair a target school with a reach school than with a safety school as long as the choices are equivalent in their actuarial value. By the elimination argument, the same result holds with respect to adding a reach school to a portfolio that is otherwise balanced.

Further analysis of the optimal application strategy provides a novel explanation of findings from regression studies of college applicant behavior and speculate about policy interventions to improve student equity and welfare. We conjecture that broad policies to decrease the costs of application may yield a greater increase in student equity than interventions that narrowly target high-achieving, low-income students.

Classical portfolio optimization models in the Markowitz \cite{markowitz1952} tradition feature risk aversion as an exogeneous \emph{parameter.} The optimal portfolio is expressed as a proportion of the total budget, and this allocation is constant for budgets of all sizes. Therefore, the value of a rational investor's risk parameter can be determined merely by locating her portfolio allocation along the efficient frontier, even if her total investment budget is unknown.

In the college application problem, the value $t_0$ of the noncollege option can function as a risk-aversion parameter. However, the analysis of the distributional heuristic reveals that regardless of their underlying risk preferences, students must modulate their risk allocation---that is, the relative share of reach and safety schools in their portfolio---according to the number of schools to which they can afford to apply. Students with a large application budget can tolerate, and indeed benefit, from taking larger risks than those without. This means that in college application, variation in students' risk allocation arises from a combination of variance in underlying preferences and variance in students' application budgets. This property, which we may call \emph{optimal risk elasticity,} is not present in the Markowitz model.

A classic finding of behavioral psychology is that agents with limited resources tend to make more risk-averse decisions than their wealthier peers, especially with respect to education and career choice \cite{hartlaubandschneider2012,vanhuizenandalessie2019}. More concretely, wealthy students tend to apply to more reach schools than low-income students of similar academic ability, a tendency that some education scholars attribute to a lack of teachers or older peers who attended selective universities \cite{hoxbyandavery2012}. But given that low-income students tend to apply to fewer schools overall, optimal risk elasticity provides a rational explanation for the difference between their risk allocation and that of their wealthy peers. Hence, the failure of low-income students to apply to reach schools is not necessarily suboptimal decisionmaking due to risk aversion, but could be \emph{optimal} decisionmaking under a smaller budget. If the latter explanation is correct, then optimal risk elasticity suggests that reducing the costs of college application for all students could improve student equity without the need to overcome psychological factors such as risk aversion and the absence of high-achieving role models.

\section{Conclusion} \label{sectionConclusion}
%

Example \ref{tightexampleforoneoverhopt}, in which $v(\mathcal{X}_h)$ is $O(h)$, demonstrates that the objective function of the college application problem approaches linearity when the admissions probabilities $f_j$ are very small. In the knapsack-constrained college application problem, Kapur \cite{kapur2022} exploits this property to reduce the problem to the binary knapsack problem and establish its NP-completeness.